---
title: "Case Studies"
subtitle: ""
author: "The R Bootcamp @ Erfurt<br/><a href='https://therbootcamp.github.io'>www.therbootcamp.com</a><br/><a href='https://twitter.com/therbootcamp'>@therbootcamp</a>"
date: "June 2018"
output:
  xaringan::moon_reader:
    css: ["default", "../_css/my-theme.css"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9'

---

layout: true

<div class="my-footer"><span>
<a href="https://therbootcamp.github.io/"><font color="#7E7E7E">Erfurt, June 2018</font></a>
&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;
&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;
&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;
&emsp;&emsp;&emsp;&emsp;&emsp;
<a href="https://therbootcamp.github.io/"><font color="#7E7E7E">www.therbootcamp.com</font></a>
</span></div> 

---

```{r, eval = FALSE, echo = FALSE}
# Code to knit slides
xaringan::inf_mr('_sessions/D1S4_CaseStudies/CaseStudies.Rmd')
```

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
# see: https://github.com/yihui/xaringan
# install.packages("xaringan")
# see: 
# https://github.com/yihui/xaringan/wiki
# https://github.com/gnab/remark/wiki/Markdown
options(width=110)
options(digits = 4)
```

# Practical - Case studies

Work on one or all of 3 case studies designed to repeat the central elements of **Intro to R**, **Data Wrangling**, and **Statistics**<br><br>

### 1 Clinical trial
Analyse a clinical trial comparing the effectiveness of drugs on reducing histamine.<br><br>

### 2 Finance
Study the development of key financial markets across the last 20 years<br><br>




